---
title: "ggplot2"
author: "Kevin.C"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggiraph)
library(dplyr)
library(patchwork) # vis composition
library(plotly) # interaction with ggplotly()

# Several datasets I used here
# mpg data - data set provided in R, all about cars
data("mpg")

# Jellybeans.dat
# This data includes the weight of jellybeans from different samples and trays
jellybeans <- read.delim("C:/Users/asus/Desktop/CU boulder/Course/S22/EMEN 5042 Data Science for quality improvement/WK8 - Process Performance Analysis/HW7/jellybeans.dat")
jellybeans$Sample <- as.factor(jellybeans$Sample)
jellybeans$Tray <- as.factor(jellybeans$Tray)

# Earing.dat
# Measurements from different lots
Earing <- read.delim("C:/Users/asus/Desktop/CU boulder/Course/S22/EMEN 5042 Data Science for quality improvement/WK13 ANOVA/hw9/Earing.dat")
Earing$lot = as.character(Earing$lot)

# HW10 data
# Height measurement for 10 parts from different operators, repeated measure for 2 times.
height <- read.delim("C:/Users/asus/Desktop/CU boulder/Course/S22/EMEN 5042 Data Science for quality improvement/wk14/height.dat")

# Some vis from another dataset
# Segment analysis for some famous cycling routes around Boulder
# complete time and power output by different cyclists.
segment <- read.csv("C:/Users/asus/Desktop/CU boulder/Course/S22/INFO 5602 information visaulization/Final Project/FinalProject_Wrapup/WriteUp_Artifact/Data/segment.csv")
segment <- na.omit(segment)
segment$w_kg_ratio_int <- as.factor(segment$w_kg_ratio_int)
```

# ggplot2 Introduction

ggplot2 is an open-source data visualization package for the statistical programming language R. Created by Hadley Wickham in 2005, ggplot2 is an implementation of Leland Wilkinson's Grammar of Graphics—a general scheme for data visualization which breaks up graphs into semantic components such as scales and layers. ggplot2 can serve as a replacement for the base graphics in R and contains a number of defaults for web and print display of common scales. Since 2005, ggplot2 has grown in use to become one of the most popular R packages. It's also the foundation of many graphic applications such as Vega-Lite and Tableau.

The grammar of graphics breaks up graphs into several components:

        -Data: 
                Defines what can be done with it.Sometimes you might need to manipulate
                data first.
        -Mapping: 
                Generic datasets to be understood by the graphic system. Aesthetic mapping
                would link variables in data to graphical properties in the geometry
        -Statistic:
                Transform input variables to displayed values
                Ex: 
                • Count number of observations in each category for a bar chart 
                • Calculate summary statistics for a boxplot. 
        -Scale:
                A scale translate back and forth between variable ranges and property ranges 
                Ex: Categories → Colour

        -Geometries:
                How to interpret aesthetics as graphical representations                
        -Facets:
                Define the number of panels with equal logic and split data among them
        -Coordinates
                Defines the physical mapping of the aesthetics to the paper
        -Theme
                Theming spans every part of the graphic that is not linked to data

With that being said, the coding will be like stacking several layers to create a highly customized visualization as below.

```{r example plot}
HL_lot_num <- c("682517","682519","682523")

Highlight <- Earing %>% 
        filter(lot %in% HL_lot_num)

# Stacking several functions to customize the vis.
 ggplot(Earing) +
        geom_boxplot(aes(lot,earing),fill="#52FF00",outlier.shape = 1,outlier.size = 2,) +
        geom_boxplot(aes(lot,earing),data = Highlight,fill="#FF3E00",
                     outlier.shape = 1,outlier.size = 2,) +
        geom_hline(aes(yintercept = 1.85),color = "red",size = 0.7)+
        geom_hline(aes(yintercept = 2.55),color = "red",size = 0.7)+
        theme_bw()+
        theme(axis.text.x = element_text(angle = 45, hjust = 1))+
        ggtitle("boxplot for lots",subtitle = "15 lots in total")+
        labs(x="Lot Number",y="Earing value")

# Interaction example - it shows tooltip when hovering on the point.
interaction_plot <- ggplot(mpg)+
  geom_point_interactive(aes(cty,hwy,color = model,tooltip = model,data_id= model))
girafe(ggobj = interaction_plot)

# "ggiraph" for interaction
# https://davidgohel.github.io/ggiraph/
```

Although it looks complicated, you can actually create a simple chart with the data by mapping them into the corresponding geometry with only one or two lines of codes. We'll make some simple chart in the following.

# Basics
ggplot2 is based on the grammar of graphics, the idea that you can build every graph from the same components: a data set, a coordinate system, and geoms—visual marks that represent data points. To display values, map variables in the data to visual 
properties of the geom (aesthetics) like size, color, and x.
and y locations.

#### * Note that we use "+" to add layers.

         ------- We have to provide <DATA>, choose <GEOM_FUNCTION> and <MAPPINGS> 
         variables you want to map-------
         
        ggplot (data = <DATA> ) +  
         <GEOM_FUNCTION> (mapping = aes( <MAPPINGS> ), 
         
         ------- Code below are not required, only use them when you want to add
         extra stuff ------
         
         stat = <STAT> , position = <POSITION> ) +  
         <COORDINATE_FUNCTION> + 
         <FACET_FUNCTION> +
         <SCALE_FUNCTION> +
         <THEME_FUNCTION>

##### We will continue to use jellybeans.dat, Earing.dat to make some plots 
 
# Boxplot / Histgram
You can define data sets and variables in ggplot() or geom_function(). Defining in the ggplot() would make it become a global setting, while the another would only effect in the certain layer.

```{r boxplot_example}
# Provide dataset, geom type and variables
ggplot(Earing) + geom_boxplot(aes(lot,earing))

# The data and mapping setting can also be defined in ggplot()
ggplot(Earing,aes(lot,earing)) + geom_boxplot()

# Also doable if you want to define all of it in geom_boxplot
ggplot() + geom_boxplot(aes(lot,earing),data = Earing)

# Hist
ggplot(jellybeans)+geom_histogram(aes(Weight))
ggplot(jellybeans)+geom_histogram(aes(Weight),color = "black", fill = "white")



# qplot - another way to use ggplot2, but I never use this one
qplot(lot,earing,data = Earing,geom = "boxplot")

# Save plot to working directory
ggsave("lot.png", width = 5, height = 5)

```

# Point/Jitter

###### A jitter plot is a variant of the strip plot with a better view of overlapping data points, used to visualise the distribution of many individual one-dimensional values.

```{r point_example}
# Provide dataset, geom type and variables
ggplot(Earing) + geom_point(aes(lot,earing),size = 2)
# Adjust width - Recommend you adjust it and see how it goes.
ggplot(Earing) + geom_jitter(aes(lot,earing),width = 0.1,size = 2)
ggplot(Earing) + geom_jitter(aes(lot,earing),width = 0.3,size = 2)

# Show it together / I changed the point type here.
plot <- ggplot(Earing) + 
  geom_boxplot(aes(lot,earing),fill = "#a1d99b",alpha = 0.5)+ 
  geom_jitter(aes(lot,earing),width = 0.1,shape = 4,color = "#3182bd");plot 
```

# Add Line & Annotation
```{r ggplot2_line}
# Line type
# http://sape.inf.usi.ch/quick-reference/ggplot2/linetype

# Horizontal Line
plot <- ggplot(Earing) + 
  geom_boxplot(aes(lot,earing),fill = "#a1d99b",alpha = 0.5)+ 
  geom_jitter(aes(lot,earing),width = 0.1,shape = 4,color = "#3182bd")+
  geom_hline(aes(yintercept = 1.85),color = "red",size = 0.7,linetype ="longdash")+
  geom_hline(aes(yintercept = 2.55),color = "red",size = 0.7,linetype ="longdash");plot

# Vertical Line
plot <- ggplot(Earing) + 
  geom_boxplot(aes(lot,earing),fill = "#a1d99b",alpha = 0.5)+ 
  geom_jitter(aes(lot,earing),width = 0.1,shape = 4,color = "#3182bd")+
  geom_hline(aes(yintercept = 1.85),color = "red",size = 0.7,linetype ="longdash")+
  geom_hline(aes(yintercept = 2.55),color = "red",size = 0.7,linetype ="longdash")+
  geom_vline(aes(xintercept = 5.5),color = "#3182bd",size = 0.5)+
  geom_vline(aes(xintercept = 11.5),color = "#3182bd",size = 0.5);plot

# Make a df for limit value
limit = c(1.85,2.55)
Limit = as.data.frame(limit)

plot <- ggplot(Earing) + 
  geom_boxplot(aes(lot,earing),fill = "#a1d99b",alpha = 0.5)+ 
  geom_jitter(aes(lot,earing),width = 0.1,shape = 4,color = "#3182bd")+
  geom_hline(data = Limit,aes(yintercept = limit),
             color = "red",size = 0.7,linetype ="longdash");plot

# Annotation
plot <- ggplot(Earing) + 
  geom_boxplot(aes(lot,earing),fill = "#a1d99b",alpha = 0.5)+ 
  geom_jitter(aes(lot,earing),width = 0.1,shape = 4,color = "#3182bd")+
  geom_hline(data = Limit,aes(yintercept = limit),
             color = "red",size = 0.7,linetype ="longdash")+
  annotate(geom = "text",x = c(1.5,1.5),y = c(1.8,2.6),
           label = c("LPL","UPL"), color ="red");plot
```

# Color
```{r ggplot2_color}
# https://colorbrewer2.org/#type=sequential&scheme=PuBuGn&n=8

# Fill vs color
# Fill
plot <- ggplot(Earing) + 
  geom_boxplot(aes(lot,earing),fill = "#a1d99b",alpha = 0.5)  ; plot
# Color
plot <- ggplot(Earing) + 
  geom_boxplot(aes(lot,earing),color = "#a1d99b",alpha = 0.5)  ; plot

# Mapping fill with data
plot <- ggplot(Earing) + 
  geom_boxplot(aes(lot,earing,fill=lot),alpha = 0.5)  ; plot

hist <- ggplot(Earing) + 
  geom_histogram(aes(earing,fill=lot),alpha = 0.7) ; hist

# Highlight specific group
# (1) specify item in mapping
plot <- ggplot(Earing) + 
  geom_boxplot(aes(lot,earing,fill = lot == 682517),alpha = 0.5) ; plot
# (2) add another layer include highlighted groups 
HL_lot_num <- c("682517","682519","682523")

Highlight <- Earing %>% 
        filter(lot %in% HL_lot_num)

plot <- ggplot(Earing) + 
  geom_boxplot(aes(lot,earing),alpha = 0.5,fill = "#a1d99b") +
  geom_boxplot(aes(lot,earing),data = Highlight, fill = "#fb9a99"); plot

# "alpha" = opacity

# scale_color/fill

# scale_color_manual() : to use custom colors
# scale_color_brewer() : to use color palettes from RColorBrewer package
# scale_color_grey() : to use grey color palettes
# Have same functions for fill
# scale_fill_brewer
# scale_fill_manual
# scale_fill_grey

# Manual; can also alter the legend through manual function
plot <- ggplot(Earing,aes(lot,earing)) + 
  geom_boxplot(aes(fill = lot))+
  scale_fill_manual(
    values = c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99'
               ,'#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a',
               '#ffff99','#b15928',"red","red","red","red","red"),
    name="test",
    labels=c("a","b")); plot

# Brewer; Since there are too many groups here, there are some boxs without fill.
plot <- ggplot(Earing,aes(lot,earing)) + 
  geom_boxplot(aes(fill = lot))+
  scale_fill_brewer(type = "qual",palette = "Accent");plot

# Palette: Check what palette included in the package
RColorBrewer::display.brewer.all()

```

# Coordinate / faceting
```{r coordinate and faceting}
# Original histogram; turn off the legend since I am going to combine 2 plots
# with the same legend.
hist <- ggplot(jellybeans) + 
  geom_histogram(aes(Weight,fill=Tray),alpha = 0.7,bins = 70)+
  theme(legend.position = "none");hist
# adjust coordinate; use it like zooming in within certain range
hist_zoom <- ggplot(jellybeans) + 
  geom_histogram(aes(Weight,fill=Tray),alpha = 0.7,bins = 70)+
  coord_cartesian(xlim=c(1.08,1.11)); hist_zoom

# Put it together
hist+hist_zoom

# Can use it with patchwork, use theme_bw for zoom plot and adjust y axis as well. 
# step 1: make some adjustment for zoom plot first
# y axis adjustment / change background / add title
hist_zoom <- ggplot(jellybeans) + 
  geom_histogram(aes(Weight,fill=Tray),alpha = 0.7,bins = 70)+
  coord_cartesian(xlim=c(1.08,1.11),ylim = c(0,90))+
  theme_bw()+
  labs(title = "Zoom in for weight 1.08 to 1.11"); hist_zoom

# Step 2 layout and composition
layout <- "
aaa##
aaa##
aaa##
aaabb
aaabb
"
# "#" means empty
hist+hist_zoom+plot_layout(design = layout)

# Some other coordinate application
# flip
hist <- ggplot(jellybeans) + 
  geom_histogram(aes(Weight,fill=Tray),alpha = 0.7,bins = 70)+
  coord_flip(); hist
# polar coordinate
hist <- ggplot(jellybeans) + 
  geom_histogram(aes(Weight,fill=Tray),alpha = 0.7,bins = 70)+
  coord_polar(); hist

# Faceting; Take 12 Tray as example
jellybeans <- read.delim("C:/Users/asus/Desktop/CU boulder/Course/S22/EMEN 5042 Data Science for quality improvement/WK8 - Process Performance Analysis/HW7/jellybeans.dat")

# jellybeans <- read.delim("~/Desktop/ggplot2/jellybeans.dat")
first12Tray <- jellybeans %>% 
        filter(Tray <= 12)
first12Tray$Tray = as.factor(first12Tray$Tray)

hist <- ggplot(first12Tray) + 
  geom_histogram(aes(Weight,fill=Tray),alpha = 0.7,bins = 70) ; hist

# facet_wrap
hist <- ggplot(first12Tray) + 
  geom_histogram(aes(Weight,fill=Tray),alpha = 0.7,bins = 70) + 
  facet_wrap(vars(Tray)); hist

# No need to use color mapping with faceting, IMO. 
# Thus I use black/white to plot bars
hist <- ggplot(first12Tray) + 
  geom_histogram(aes(Weight),fill = "white",color = "black",bins = 70) + 
  facet_wrap(vars(Tray)); hist

```

# Title/Labels
``` {r title and label}
# labs()
hist+
  labs(x = "WEIGHT",y= "COUNT",
      title = "Histograms by Tray #",
      subtitle = "- First 12 Tray",)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

# Theme
```{r theme}
# Adjust background
t <- plot <- ggplot(Earing) + 
  geom_boxplot(aes(lot,earing),fill = "#a1d99b",alpha = 0.5)+ 
  geom_jitter(aes(lot,earing),width = 0.1,shape = 4,color = "#3182bd")+
  geom_hline(aes(yintercept = 1.85),color = "red",size = 0.7,linetype ="longdash")+
  geom_hline(aes(yintercept = 2.55),color = "red",size = 0.7,linetype ="longdash");t 

t + theme_bw()

t + theme_dark()

# rotate x-axis label
# There are many setting can be manipulate in the theme()
# try ?theme() for the document
t + theme(axis.text.x = element_text(angle = 45, hjust = 1) )
```

# Interaction
```{r interaction}
# interaction
lot_boxplot <- t + theme(axis.text.x = element_text(angle = 45, hjust = 1) )
ggplotly(lot_boxplot)


jellybeans$Sample <- as.factor(jellybeans$Sample)
jellybeans$Tray <- as.factor(jellybeans$Tray)


hist <- ggplot(jellybeans) + 
  geom_histogram(aes(Weight,fill=Tray),alpha = 0.7, position = "identity",bins = 70)+
  geom_vline(xintercept = c(1.09,1.13),color = "red" , linetype = "longdash",size =1); 

ggplotly(hist)

```

# HW 10 Vis

```{r hw10}
height$Part <- as.factor(height$Part)
height$Operator <- as.factor(height$Operator)
height$Repetition <- as.factor(height$Repetition)

a <- ggplot(data = height,aes(Part,Value))

p1 <-  a + geom_boxplot(fill = "#b2df8a");p1

p2 <-  a + 
  geom_point(aes(color = Operator),size = 3) + 
  stat_summary(fun = "mean",geom="point",shape =4,size =3);p2

# Faceting
p3 <-  a + 
  geom_point(aes(color = Operator),size = 3) + 
  facet_grid(cols=vars(Operator));p3

# Install package first
p1+p2  

# define layout
layout <- "
aaaccc
aaaccc
aaaccc
bbbccc
"
p4 <- ggplot(data = height,aes(Operator,Value))+
  geom_boxplot(fill = "#ffed6f");p4

p1+p4+p2+plot_layout(design = layout)

```

# Some other visualization - Strava segments analysis. 

I collect results of 3 segments from 1700 riders thru strava api and visualize the data to see if they are related. Also got the power-to-weight ratio (Higher means stronger) and used them mapping with color. I am trying to use this plot to predict the results. Like if someone can finish segment A in 20 mins, he might make it around 45 mins for segment B.
```{r other example}
"  Variables:
segment$rider
segment$SunshineHillclimbChallengeCourse_sec
segment$FourmiletoGoldHill_sec
segment$SuperFlag_sec
segment$w_kg_ratio
segment$SunshineHillclimbChallengeCourse_min
segment$FourmiletoGoldHill_min
segment$SuperFlag_min
"
# point
ggplot(segment)+
  geom_point(aes(SuperFlag_min,FourmiletoGoldHill_min,color = w_kg_ratio_int))+
    theme_minimal()

# boxplot 
ggplot(segment)+
  geom_point(aes(SuperFlag_min,FourmiletoGoldHill_min,color = w_kg_ratio_int))+
  geom_boxplot(aes(SuperFlag_min,FourmiletoGoldHill_min,group = SuperFlag_min),fill = "#74c476")+
  theme_minimal()

# interaction
interaction_plot <- ggplot(segment)+ 
  geom_point_interactive(aes(SuperFlag_min,FourmiletoGoldHill_min,color=w_kg_ratio_int,
                             tooltip =rider,data_id=w_kg_ratio_int))+
  scale_color_brewer(type = "qual",palette = "Accent") 

# Highlight by w/kg group
test_p <- ggplot(segment)+
  geom_point(aes(SuperFlag_min,FourmiletoGoldHill_min,color=w_kg_ratio_int),size = 1.5)

# filter 
ggplotly(test_p)

# interaction, hovering
girafe(ggobj=interaction_plot)
```


# Reference:

 - Manual: 
 
        https://cran.microsoft.com/snapshot/2015-01-06/web/packages/ggplot2/ggplot2.pdf

 - sorting:
 
         https://www.statology.org/order-bars-ggplot2-bar-chart/#:~:text=By%20default%2C%20ggplot2%20bar%20charts%20order%20the%20bars,ordering%20the%20bars%20in%20some%20other%20specific%20order.
 
 - Color brewer:
 
        https://colorbrewer2.org/#type=sequential&scheme=PuBuGn&n=8
        
 -Palette
      
          > RColorBrewer::display.brewer.all()
 
 - Scale color :
 
        https://ggplot2.tidyverse.org/reference/scale_brewer.html#palettes
 
        * type	= One of seq (sequential), div (diverging) or qual (qualitative)
       
 - Interaction
 
        - package: plotly
       
 - Animation 
 
        - packages: gganimate / gifski / av